---
title: "Barnard's test template"
author: "Emily Hatch"
date: "2023-08-22"
output: 
  html_document:
   keep_md: TRUE

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = TRUE, warning = FALSE, message = FALSE, error = FALSE) 
```

#### Objective
Statistical analysis of 2x2 tables in R

#### Approach 1: Compare all samples to a control sample (Amanda's code)
Input is a dataframe of counts.

```{r all versus 1}
library(Barnard)
library(tidyverse)

df_1 <- data.frame(condition = c('DMSO', 'BLEB', 'B5', 'B6', 'B7_2uM', 'B8', 'D9', 'B7_500nM'), 
     NOTRUP = c(155, 155, 122, 163, 175, 211, 130, 168),
     RUP = c(55, 28, 38, 47, 46, 40, 58, 38))

ctrl <- df_1 %>% filter(grepl('DMSO', condition))

NEGctrl <- ctrl$NOTRUP # output is a single named value
POSctrl <- ctrl$RUP # output is a single named value

stats <- categories %>%
     rowwise() %>%
     mutate('p' = barnard.test(NEGctrl, NOTRUP, POSctrl, RUP)$p.value[2]
          ) %>% # p.value[2] specifies which value in the output to report in the 'p' column
     mutate('significant' = case_when(
          p <= 0.05 ~ '*')) # adds column of p value text
```
#### Approach 2: All versus all pairwise comparisons.
This uses a package from the Fred Hutch Vaccine Immunology Statistical Center. https://rdrr.io/github/FredHutch/VISCfunctions/man/pairwise_test_bin.html

In this approach, the data is transformed into a case table and then analyzed by chisquare.

This data is further simplified into two vectors: sample id and logical value.

After the significance of all possible combinations are calculated, results are filtered to those of interest and a post-test is performed.

Example input: dataframe of counts, with ColNames = Condition, Intact, Rupture. 
```{r all versus all}
install.packages("remotes")
remotes::install_github("FredHutch/VISCfunctions", build_vignettes = TRUE)
library(tidyverse)
library(VISCfunctions)
library(rstatix)
```
Generate contingency table for chi-square analysis from count dataframe df_1

```{r all versus all}
# generate a contingency table from dataframe with pooled values. This is used as input to chi square
table_pooled <- df_1 %>%
  column_to_rownames(var = "condition") %>%
  select(NOTRUP, RUP) %>%
  as.matrix()

#Add category names to rows and columns of contingency table.
condition <- rownames(table_pooled)  
dimnames(table_pooled) = list("Condition" = condition, "Rupture" = c("NOTRUP", "RUP"))

str(table_pooled)
```
Run chi square
```{r all versus all}
chisq.test(table_pooled)
```
Convert contingency table to vectors for multiple Barnard's test.
```{r all versus all}
df_cases <- table_pooled %>%
  counts_to_cases()

#convert case factor variable to binary numbers
df_cases$Rupture <- as.character(df_cases$Rupture) # Necessary if some factors were used
df_cases$Rupture[(df_cases$Rupture == "NOTRUP")] <- 0 # Take the subset of values in column "Rupture" of dataframe df_cases that equal "NOTRUP" and replace them with the value "0".
df_cases$Rupture[(df_cases$Rupture == "RUP")] <- 1 
df_cases$Rupture <- as.numeric(df_cases$Rupture)

# Run barnard's test after modifying the groups to be vectors
B_pair <- df_cases %>%
  group_modify(~ as.data.frame(
       pairwise_test_bin(x = .$Rupture, group = .$Condition)))

B_pair
```

```{r all versus all}
# add a text p-value column
B_pair <- B_pair %>% 
  add_significance(p.col = "ResponseTest", output.col = "p_value") 

# filter for useful comparisons
B_pair_s <- B_pair %>%
  filter(Comparison == "DMSO vs. BLEB" | Comparison == "DMSO vs. B5" | Comparison == "DMSO vs. B8")
```

Perform post-test
```{r all versus all}
# Split "Comparison" column into group 1 and 2 to calculate adjusted pvalue
B_pair_sigf <- B_pair_s %>%
  separate_wider_delim(cols = Comparison, delim = " vs. ", names = c("group1", "group2")) %>%
  adjust_pvalue(p.col = "ResponseTest", output.col = "p.adjust", method = "BH") %>%
  add_significance(p.col = "p.adjust", output.col = "p.adjust.value") #%>%
  mutate(y.position = c(1, 1.1, 0.9, 0.95, 0.8, 0.7)) # manually add y positions for graphing

B_pair_sigf

```

