---
title: "Barnard's test template"
author: "Emily Hatch"
date: "2023-08-22"
output: 
  html_document:
   keep_md: TRUE

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = TRUE, warning = FALSE, message = FALSE, error = FALSE, fig_width = 4, fig_height = 6, fig_retina = TRUE, dev = c("png", "pdf")) 
```

#### Objective
Add the results of Barnard’s test to a dataframe in R

#### Pseudocode
load the necessary libraries
create a dataframe
identify the control values for comparison
use rowwise() and mutate() to create a column for the results
format to call the correct results barnard.test()$p.value[2]
Note that this will call the two-sided p-value, if you would like to switch to a 1-sided test you can change it to say barnard.test()$p.value[1]

```{r dependencies}
library(Barnard)
library(tidyverse)

categories <- data.frame(condition = c('DMSO', 'BLEB', 'B5', 'B6', 'B7_2uM', 'B8', 'D9', 'B7_500nM'), 
     NOTRUP = c(155, 155, 122, 163, 175, 211, 130, 168),
     RUP = c(55, 28, 38, 47, 46, 40, 58, 38))

ctrl <- categories %>% filter(grepl('DMSO', condition))

NEGctrl <- ctrl$NOTRUP
POSctrl <- ctrl$RUP

stats <- categories %>%
     rowwise() %>%
     mutate('p' = barnard.test(NEGctrl, NOTRUP, POSctrl, RUP)$p.value[2]
          ) %>%
     mutate('significant' = case_when(
          p <= 0.05 ~ '*'))
```

I also added a column that automatically adds a star to data that is significant so you could add this to a graph easily using

geom_text(data = stats, aes(x = condition, y = 0.6, label = paste(significant)))

Note that for a large table this is going to take a good chunk of time to run, so you may need to be a bit patient.